---
title: "Lecture 9 hands on"
author: "Colin McSweeney"
date: "May 1, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
url <- "https://bioboot.github.io/bimm143_W18/class-material/WisconsinCancer.csv"

# Complete the following code to input the data and store as wisc.df
wisc.df <- read.csv(url)
head(wisc.df)
```



How many M's do we have?
```{r}
table(wisc.df$diagnosis)
```

Use Logicals- make a numeric classification

```{r}
diagnosis <- as.numeric(wisc.df$diagnosis == "M")
table(diagnosis)
```

```{r}
# Convert the features of the data: wisc.data
wisc.data <- as.matrix(wisc.df[,3:ncol(wisc.df)])
rownames(wisc.data) <- wisc.df$id
head(wisc.data)
wisc.data <- wisc.data[, -31]
head(wisc.data)
```




**Q1.**, How many observations are in this dataset?

```{r}
nrow(wisc.data)
```

**Q2**. How many variables/features in the data are suffixed with _mean?
```{r}
#return things with the word 'mean' in them
grep("mean", colnames(wisc.data), value = TRUE)

#return the position, or index, of things with 'mean' in them
grep("mean", colnames(wisc.data), value = FALSE) #default

#return things that dont have the word 'mean' in them
grep("mean", colnames(wisc.data), value = TRUE, invert = TRUE)

```

**Q3**. How many of the observations have a malignant diagnosis?
```{r}

```




Part 2

```{r}
# Check column means and standard deviations
colMeans(wisc.data)

apply(wisc.data,2,sd)
```

```{r}
# Perform PCA on wisc.data by completing the following code
wisc.pr <- prcomp(wisc.data, scale = TRUE)
```

```{r}
# Look at summary of results
summary(wisc.pr)
```

```{r}
plot(wisc.pr)
```

**Q4**. From your results, what proportion of the original variance is captured by the first principal components (PC1)?
**Q5**. How many principal components (PCs) are required to describe at least 70% of the original variance in the data?
**Q6**. How many principal components (PCs) are required to describe at least 90% of the original variance in the data?

```{r}
biplot(wisc.pr)
```

# Scatter plot observations by components 1 and 2
```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col = diagnosis +1,
     xlab = "PC1", ylab = "PC2")
```



Scree varianc
```{r}
pr.var <- wisc.pr$sdev^2

#portion of varianc
pve <- pr.var / sum(pr.var)

#plot variance explained for each principal componenet
plot(pve, xlab = "Principal Component",
     ylab = "Propotion of vriance explaines",
     ylim = c(0, 1), type = "o")
```






NEXT PART

#Hierarchical CLustering

```{r}
#scale the wisc.data 
data.scaled <- scale(wisc.data)

#Calculate distance matrix needed for hclust
data.dist <- dist(data.scaled)

#
wisc.hclust <- hclust(data.dist)
```

Plot our hclust model tree

```{r}
plot(wisc.hclust)
abline(h=20, col = "red")
wisc.hclust.clusters <- cutree(wisc.hclust, h= 20)
```


```{r}
table(wisc.hclust.clusters, diagnosis)
```



Section4 

```{r}
wisc.km <- kmeans(scale(wisc.data), centers=2, nstart= 20)
table(wisc.km$cluster, diagnosis)
```
comparing kmeans with hclust

```{r}
table(wisc.hclust.clusters, wisc.km$cluster)
```

Part 5


```{r}
## Use the distance along the first 7 PCs for clustering i.e. wisc.pr$x[, 1:7]
wisc.pr.hclust <- hclust(dist(wisc.pr$x[, 1:7]))
plot(wisc.pr.hclust)
```

```{r}
wisc.pr.hclust.clusters <- cutree(wisc.pr.hclust, k=4)
```

```{r}
table(wisc.pr.hclust.clusters, diagnosis)
```

