---
title: "Bioinformatics Class 7"
author: "Colin McSweeney"
date: "April 24, 2018"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Functions Again
Here I am going to revisit our function from class 6

```{r}
source("http://tinyurl.com/rescale-R")
```


lets see if we can use this function

```{r}
rescale(1:10)
```

looks good let's break it!

```{r}
#rescale(c(1:10, "string"))
```

lets try the **rescale2()** function

```{r}
#rescale(c(1:10, "string"))
```

##Write an NA checking function

Here we write a new function to check for NAs in two inputs 

```{r}
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)
```


```{r}
sum(is.na(x))
```

Let's put this together with x and y

```{r}
is.na(x)
is.na(y)
```


```{r}
sum(is.na(x) & is.na(y))
```


This is the both_na starting point

```{r}
both_na <- function(x, y) {
  ## Check for 
 sum( is.na(x) & is.na(y) )
}
```



Test

```{r}
both_na(x,y)
```

and again

```{r}
x <- c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)

both_na(x,y2)

```

```{r}
both_na2 <- function(x, y) {
 if(length(x) != length(y)) {
 stop("Input x and y should be the same length")
 }
 sum( is.na(x) & is.na(y) )
}
```


```{r}
both_na2(x,y2)
```



try both_na3() fuunction

```{r}
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)

ans <- both_na3(x,y)
```

```{r}
ans
```

##another function example: gene intersection

```{r}
df1
df2

x <-df1$IDs
y <- df2$IDs
```

```{r}
x
y
```

```{r}
#intersect(x,y)

x%in% y
```

```{r}
#intersection(y,x)

y %in% x
```

Now we can access the genes we want with these 'indices'
```{r}
x[x %in% y]
y[y %in% x]
```

can make these columns of the same object by using cbind()

```{r}
cbind(x[x %in% y], y[y %in% x])
```

```{r}
gene_intersect2(df1, df2)
```
Lets try the merge () function for this job

```{r}
merge(df1, df2, by= "IDs")
```

